---
title: "DataCleaning2017"
author: "Anne"
date: "06/05/2020"
output:
  html_document:
    code_folding: hide
    fig_height: 5
    fig_width: 5
    number_sections: yes
    theme: united
    toc: yes
  pdf_document:
    toc: yes
---
```{r}
library(readxl)
library(dplyr)


patients <- read.csv("C:/Users/Anne/Dropbox (Sydney Uni Student)/SallyChen/PaedWork/42510_paeds_allocation_project/patients.csv", na.strings=c("-" ," ", "","NA"))

donor <- read.csv("C:/Users/Anne/Dropbox (Sydney Uni Student)/SallyChen/PaedWork/42510_paeds_allocation_project/donor.csv",na.strings=c("-" ," ", "","NA"))

comorbidity <- read.csv("C:/Users/Anne/Dropbox (Sydney Uni Student)/SallyChen/PaedWork/42510_paeds_allocation_project/comorbidity.csv", na.strings=c("-" ," ", "","NA"))

transplants <- read.csv("C:/Users/Anne/Dropbox (Sydney Uni Student)/SallyChen/PaedWork/42510_paeds_allocation_project/transplants.csv", na.strings=c("-" ," ", "","NA"))

donor_state <- read_excel("C:/Users/Anne/Dropbox (Sydney Uni Student)/SallyChen/PaedWork/transplants_donor_state.xlsx")

# All the other information in this file `transplant_donor_state` is also contained in the `donor` file
donor_state <- donor_state[,c("id","transplantid","donorcentrestate")] 
donor_state[donor_state=="-"] <- NA

waitlist <- read.csv("C:/Users/Anne/Dropbox (Sydney Uni Student)/SallyChen/PaedWork/42510_paeds_allocation_project/waitlist.csv",na.strings=c("-" ," ", "","NA"))

```

```{r, eval=FALSE}
drugdose <- read.csv("42510_paeds_allocation_project/drugdose.csv", na.strings=c("","NA"))


rejections <- read.csv("42510_paeds_allocation_project/rejections.csv", na.strings=c("","NA"))

therapy <- read.csv("42510_paeds_allocation_project/therapy.csv", na.strings=c("","NA"))

treatments <- read.csv("42510_paeds_allocation_project/treatments.csv", na.strings=c("","NA"))

```

# Adding the Donor States + combining the donor and recipient (patient) data 

```{r}
#colnames(donor)
#colnames(patients)
#colnames(donor_state)
dim(donor_state)
print(paste("donor=",dim(donor)))
data_d<- merge(donor_state, donor, by=c("id","transplantid"))
data_rd <- merge(data_d, patients, by = "id")
dim(data_d)
print(paste("data_rd=",dim(data_rd)))


```

# Adding the transplant data 

```{r}
#colnames(transplants)
#colnames(data_rd)
print(paste("transplants=",dim(transplants)))
data_rdtx <- merge(data_rd, transplants, by= c("id",  "transplantid", "graftno", "transplantdate", "donorsourcecode"))
data_rdtx <- data_rdtx %>% 
  mutate(transplantdate = as.Date(transplantdate,  format="%d%b%Y"),
        rrtstartdate = as.Date(rrtstartdate, format="%d%b%Y")) %>% 
  mutate(recip_waittime = transplantdate - rrtstartdate)
          
          
print(paste("data_rdtx=",dim(data_rdtx)))

table(table(data_rdtx$ddonid))
length(data_rdtx$ddonid)


```


# Add the commorbidity data

```{r}
#colnames(comorbidity)
#colnames(data_rdtx)
print(paste("comorbidity=",dim(comorbidity)))
#there are 30 recordings for each patient - we take the recordings for when max(codate < tx_date)

data_rdtxc <- merge(data_rdtx, comorbidity, by= c("id"), all.x= TRUE)
data_rdtxc <- data_rdtxc %>% 
  mutate(transplantdate = as.Date(transplantdate, format="%d%b%Y"),
                                  codate = as.Date(codate, format="%d%b%Y")) 


data_rdtxc <- data_rdtxc %>% group_by(id, graftno, transplantdate) %>% 
    mutate(diff = transplantdate - codate) %>% filter(diff>=0) %>% group_by(id, graftno) %>% filter(diff == min(diff)) %>%mutate(recip_waittime = as.numeric(as.character(recip_waittime)))

print(paste("data_rdtxc=",dim(data_rdtxc)))



```

# Adding the waitlist data

```{r}
waitlist <- read.csv("C:/Users/Anne/Dropbox (Sydney Uni Student)/SallyChen/PaedWork/42510_paeds_allocation_project/waitlist.csv",na.strings=c("-" ," ", "","NA"))
# waitlist$waitorgans = factor(waitlist$waitorgans, levels=c(levels(waitlist$waitorgans), "kidneys"))
# waitlist$waitorgans[is.na(waitlist$waitorgans)] <- "kidneys"

waitlist_data <- waitlist %>%dplyr::select(id, waitorgans, waitdate, waitstatus) %>% filter(waitstatus == "Active")

# Extracting the first "Active" date for the single transplant recipients 
waitdata <- waitlist_data[!duplicated(waitlist_data$id), ]

data_rdtxcw <- merge(data_rdtxc,waitdata, by = "id" ) %>% mutate(waitdate = lubridate::dmy(waitdate), recip_listtime = as.numeric(transplantdate - waitdate))

print(paste("data_rdtxcw=",dim(data_rdtxcw)))



```


# Extracting Deceased Donors, (keeping multiple transplant) and transplant year 2008+

This is done twice, once for the data set and ones for the waitlist

```{r }
data_rdtxcw_final <- data_rdtxcw  %>%
  filter(donorsourcecode == "Deceased", 
          transplantdate > "2008-09-18", 
         is.na(waitorgans),
         waitdate > "2006-06-28", 
         rrtstartdate > "2006-06-28"
         )%>%
  as.data.frame()
data_rdtxcw_final <- data_rdtxcw_final %>% mutate_if(is.factor, as.character) %>% mutate(id = as.character(id)) 

data_rdtxcw_final$waitorgans <- "kidneys"

print(paste("data_rdtxcw_final=",dim(data_rdtxcw_final)))


```


# Calculating epts and kdri

```{r}

index1 <- which(is.na(data_rdtxcw$ageattransplant) |
                             is.na(data_rdtxcw$diabetescode_text) | 
                             is.na(data_rdtxcw$recip_waittime))  

sub_data <- data_rdtxcw[-index1,]
length(index1)
nrow(sub_data)

data_rdtxcw[-index1,"recip_epts"]$recip_epts <-(0.047*as.numeric(sub_data$ageattransplant>25)*(sub_data$ageattransplant-25)
   -0.015*as.numeric(!(sub_data$diabetescode_text %in% c("No")))*as.numeric(sub_data$ageattransplant>25)*(sub_data$ageattransplant-25)
                  +0.398*(ifelse(sub_data$graftno == 1, 0, 1)) 
                  -0.237*as.numeric(!(sub_data$diabetescode_text %in% c("No")))*(ifelse(sub_data$graftno == 1, 0, 1)) 
                  +0.315*log(sub_data$recip_waittime/365.25+1) 
                  -0.099*as.numeric(!(sub_data$diabetescode_text %in% c("No")))*log(sub_data$recip_waittime/365.25+1)
                  +0.130*as.numeric(sub_data$recip_waittime<365.25)
                  -0.348*as.numeric(!(sub_data$diabetescode_text %in% c("No")))*as.numeric(sub_data$recip_waittime<365.25)
                  +1.262*as.numeric(!(sub_data$diabetescode_text %in% c("No"))))



index <- which(is.na(data_rdtxcw$donorage)|
                             is.na(data_rdtxcw$donor_height)|
                             is.na(data_rdtxcw$donor_weight)|
                             is.na(data_rdtxcw$donor_hypertensioncode) |
                             data_rdtxcw$donor_hypertensioncode == "Unknown"|
                             is.na(data_rdtxcw$donor_diabetescode) | 
                             is.na(data_rdtxcw$ddon_creatinineterminal) |
                             is.na(data_rdtxcw$ddon_deathcategorycode) | 
                            is.na(data_rdtxcw$ddon_heartbeatingcode))

sub_data <- data_rdtxcw[-index,]
length(index)
nrow(sub_data)


data_rdtxcw[-index,"donor_kdri"]$donor_kdri <- exp(0.0128*(sub_data$donorage-40)
                           -0.0194*as.numeric(sub_data$donorage<18)*(sub_data$donorage-18)
                           + 0.0107*as.numeric(sub_data$donorage>50)*(sub_data$donorage-50)
                           -0.0464*(sub_data$donor_height-170)/10
                           -0.0199*as.numeric(sub_data$donor_weight<80)*(sub_data$donor_weight-80)/5
                           +0.1260*as.numeric(sub_data$donor_hypertensioncode=="Yes")
                           +0.1300*as.numeric(!(sub_data$donor_diabetescode %in% c("No", "No Diabetes")))
                           +0.0881*as.numeric(sub_data$ddon_deathcategorycode %in% c("Cerebral Infarct", "Intracranial Haemorrhage"))
                           +0.2200*(sub_data$ddon_creatinineterminal/88.42-1)
      -0.2090*as.numeric(sub_data$ddon_creatinineterminal/88.42>1.5)*(sub_data$ddon_creatinineterminal/88.42-1.5)
      +0.1330*as.numeric(sub_data$ddon_heartbeatingcode == "N"))



  
summary(data_rdtxcw$donor_kdri)
summary(data_rdtxcw$recip_epts)


```



# Add Variable names 

```{r}

#colnames(data_rdtxcw_final)

# Add the variable names
data_name_dictionary <- read_excel("data_name_dictionary.xlsx")
index <- which(is.na(data_name_dictionary$new_name))

names(data_rdtxcw)[names(data_rdtxcw) %in% c(data_name_dictionary$raw_data_name[-index])] <- c(data_name_dictionary$new_name[-index])


```

# Fixing the recip_liststartdate for multiple deceased donor transplants

```{r}
# Extracting the second/third "Active" date for multiple transplant 

multi<-names(which(table(data_rdtxcw$recip_id) == 2))

length(multi)

multi_transplant <-data_rdtxcw[data_rdtxcw$recip_id %in% multi,] %>% dplyr::select(recip_id,tx_date, recip_graftno, recip_liststartdate) %>% group_by(recip_id) %>% arrange(recip_graftno, .by_group = TRUE)

waitdata <- waitlist_data[waitlist_data$id %in% multi,] %>% mutate(waitdate = lubridate::dmy(waitdate))


waitdata <- merge(multi_transplant, waitdata, by.x = "recip_id", by.y = "id") %>% 
  mutate(diff = as.numeric(waitdate - tx_date)) %>% 
  group_by(recip_id) %>% 
  mutate(chosen = ifelse(diff >0, "eligible", "."))%>% 
  filter(chosen =="eligible") %>% 
  group_by(recip_id) %>% 
  mutate(newdate = min(waitdate))

for(i in multi){
  data_rdtxcw[data_rdtxcw$recip_id == i & data_rdtxcw$recip_graftno ==2, "recip_liststartdate"] <- waitdata[waitdata$recip_id ==i, ]$newdate[1]
}

dim(data_rdtxcw)

index<- which(is.na(data_rdtxcw$tx_organ))
data_rdtxcw$tx_organ[index] <- "kidneys"

saveRDS(data_rdtxcw, file = "data_2017_full.rds")

```


# Extracting Deceased Donors, (keeping multiple transplant) and transplant year 2008+

```{r }
data_rdtxcw_final <- data_rdtxcw  %>%
  filter(donor_source == "Deceased", 
          tx_date > "2008-09-18", 
         is.na(tx_organ),
         recip_liststartdate > "2006-06-28", 
         recip_rrtstartdate > "2006-06-28"
         )%>%
  as.data.frame()
data_rdtxcw_final <- data_rdtxcw_final %>% mutate_if(is.factor, as.character) %>% mutate(recip_id = as.character(recip_id)) 
data_rdtxcw_final$tx_organ <- "kidneys"

print(paste("data_rdtxcw_final=",dim(data_rdtxcw_final)))
saveRDS(data_rdtxcw_final, "data_2017.rds")

multi_sub<-names(which(table(data_rdtxcw_final$recip_id) == 2))
length(multi_sub)
```

# Extracting the waitlist dynamics 

```{r}
# Extract the first active case
active <- waitlist %>% 
    filter(waitstatus == "Active", is.na(waitorgans)) %>% 
    group_by(id) %>% 
    filter(waitseq == min(waitseq))

# extract all the off list cases 
off <- waitlist %>% 
    filter(waitstatus == "Off List", is.na(waitorgans)) 

#note that some people never off the last after the active date
dynamics = merge(active, off, by = c("id", "waitstat", "waitorgans"), all.x = TRUE) 
dim(dynamics)
dynamics <- dynamics %>% group_by(id) %>% mutate(waitseq.x = as.numeric(waitseq.x),
waitseq.y = as.numeric(waitseq.y), diff = waitseq.y - waitseq.x) %>% ungroup()%>% mutate(id = as.character(id)) %>% as.data.frame()
dim(dynamics)

index <- which(dynamics$diff < 0)
dynamics[index,7:9] <- NA
length(unique(dynamics$id))

#extracting the people who were never off 
no_off <- rbind(dynamics[which(is.na(dynamics$diff)),], dynamics[which(dynamics$diff %in% c(1)),] ) #never off 


############# SINGLE RECORD ########### from not NA 
check_dt <- dynamics[which(!(dynamics$id %in% no_off$id)),]
check_dt3 <- check_dt[which(check_dt$id %in% names(which(table(check_dt$id) ==1))),] #one record, imputed censored/ dead (survival)


######## MULTIPLE RECORD ########### check2 and check3 are mutually exclusive
check_dt2 <- check_dt[which(check_dt$id %in% names(which(table(check_dt$id) >1))),] #multiple record
check_dt4 <- check_dt2[check_dt2$diff >0, ] # clean multiple record
check_dt5 <- check_dt4[which(check_dt4$id %in% names(which(table(check_dt4$id) ==1))),] #multiple, imputed censored/dead (survival)

check_dt6 <- check_dt4[which(!(check_dt4$id %in% names(which(table(check_dt4$id) ==1)))),] # extract unique
check_dt6 <- check_dt6 %>% group_by(id) %>% filter(waitseq.y == min(waitseq.y)) %>% as.data.frame()

ha <- rbind(no_off, check_dt3, check_dt5, check_dt6)
dim(ha)
length(unique(ha$id))

############################################

miss<-dynamics[which(!(dynamics$id %in% ha$id)),] 
miss <- miss[!duplicated(miss$id),]
dynamics <- rbind(ha, miss)
dim(dynamics)
length(unique(dynamics$id))
 #for survival

dynamics<- dynamics%>%  
    mutate(waitdate.x = lubridate::dmy(waitdate.x), 
           waitdate.y = lubridate::dmy(waitdate.y))

index <- which(is.na(dynamics$waitdate.y))
dynamics$waitdate.y[index] <- as.character("2018-01-01")
dynamics$waitstatus.y <- as.character(dynamics$waitstatus.y)
dynamics$waitstatus.y[index] <- c("imputed censored")

dynamics <- dynamics %>% mutate(time = waitdate.y - waitdate.x)

saveRDS(dynamics, file = "data_2017dynamics_new.rds")


d1 <- dynamics[,c(1:6, 11)]
colnames(d1) <- c("id", "waitstat", "waitorgans", "waitseq", "waitdate", "waitstatus", "time")

d2<- dynamics[,c(1:3,7:9,11)]
colnames(d2) <- c("id", "waitstat", "waitorgans", "waitseq", "waitdate", "waitstatus", "time")
dynamics <- rbind(d1, d2)
dim(dynamics)

dynamics_c = merge(comorbidity, dynamics, by = c("id")) %>% 
    mutate(codate = as.Date(codate, format="%d%b%Y"))%>% tidyr::drop_na(waitstatus)
dim(dynamics_c)

class(dynamics_c$waitdate)
class(dynamics_c$codate)

dynamics_wl <- dynamics_c %>% mutate(diff = waitdate - codate )%>% filter(diff > 0) %>% group_by(id,waitstatus) %>% filter(diff == min(diff)) #changed this 30/09/2020, before it was slice(which.max)

dim(dynamics_wl)

waitlist_dyn = merge(patients,dynamics_wl, by = "id")
dim(waitlist_dyn)

waitlist_dyn$waitorgans <- "kidneys"


saveRDS(waitlist_dyn, file = "data_2017dynamics.rds")

```

```{r}
colnames(data_rdtxcw)

merged <- merge(data_rdtxcw[,-c(101, 102, 103, 104, 105)], dynamics, by.x = c("recip_id"), by.y = "id", all.x = TRUE)
summary(as.factor(merged$waitstatus.y))

```
