---
title: "DataCleaning2017"
author: "Anne"
date: "06/05/2020"
output:
  html_document:
    code_folding: hide
    fig_height: 5
    fig_width: 5
    number_sections: yes
    theme: united
    toc: yes
  pdf_document:
    toc: yes
---
```{r}
library(readxl)
library(dplyr)

## The dataset patient consists of 58232 rows and 24 columns where each row represent all patient in the registry 
# (note: includes patients on and off the waiting list) and each column represent demographical 
# and clinical variables associated with the patient. 
patients <- read.csv("./42510_paeds_allocation_project/patients.csv", na.strings=c("-" ," ", "","NA"))
dim(patients)



## The dataset donor consists of 14973 rows and 42 columns where each row represent all donors of kidneys 
# (deceased and living) and the associated graft number of kidneys donated by the donor for each recipient and 
# each column represent demographical and clinical variables associated with the donor
donor <- read.csv("./42510_paeds_allocation_project/donor.csv",na.strings=c("-" ," ", "","NA"))
dim(donor)
## The dataset donor_state consists of 14973 rows and 30 columns where each row and column are like the donor 
# dataset i.e. All other information in`transplant_donor_state` is also contained in the `donor` file
donor_state <- read_excel("./42510_paeds_allocation_project/transplants_donor_state.xlsx")
dim(donor_state)
donor_state <- donor_state[,c("id","transplantid","donorcentrestate")] 
donor_state[donor_state=="-"] <- NA



## The dataset transplants consists of 14973 rows and 29 columns where each row represent a recipient to a kidney 
# transplant and each column represent clinical variables related to the transplant
transplants <- read.csv("./42510_paeds_allocation_project/transplants.csv", na.strings=c("-" ," ", "","NA"))
dim(transplants)


## The dataset comorbidity consists of 447671 rows and 10 columns where each row represent every recording of 
# comorbidity for each patient and each column represent the comorbidity and date recorded
comorbidity <- read.csv("./42510_paeds_allocation_project/comorbidity.csv", na.strings=c("-" ," ", "","NA"))
dim(comorbidity)





## The dataset waitlist consists of 41765 rows and 6 columns where each row represent the waitlist details 
# for each patient and each column represent the waitlist details
waitlist <- read.csv("./42510_paeds_allocation_project/waitlist.csv",na.strings=c("-" ," ", "","NA"))
dim(waitlist)



## Brief look at the structure and variables in the data 
colnames(donor)
colnames(patients)
colnames(donor_state)
dim(donor_state)
print(paste("donor=",dim(donor))) 

```


# Adding the donor states and combining the donor and recipient (patient) data 

```{r}
# Merge the donor states dataset to the donor dataset
data_d<- merge(donor_state, donor, by=c("id","transplantid"))
dim(data_d) #14973 x 43

# Merge the donor and recipient dataset
data_rd <- merge(data_d, patients, by = "id")
print(paste("data_rd=",dim(data_rd))) ## 14973 x 66


```

# Adding the transplant data 

```{r}
print(paste("transplants=",dim(transplants))) # 14973 x 29

# Merge transplant data to the donor-recipient data
data_rdtx <- merge(data_rd, transplants, by= c("id",  "transplantid", "graftno", "transplantdate", "donorsourcecode"))
data_rdtx <- data_rdtx %>% 
  mutate(transplantdate = as.Date(transplantdate,  format="%d%b%Y"),
        rrtstartdate = lubridate::dmy(rrtstartdate)) %>% 
  mutate(recip_waittime = transplantdate - rrtstartdate) # calculate the recipient waittime
          
          
print(paste("data_rdtx=",dim(data_rdtx))) # 14973 x 91

table(table(data_rdtx$ddonid)) # 1545 single kidney donors and 3877 double kidney donors



```


# Add the commorbidity data

```{r}
#colnames(comorbidity)
#colnames(data_rdtx)
print(paste("comorbidity=",dim(comorbidity))) # 447671 x 10

# Merge comorbidity data to the dataset 
data_rdtxc <- merge(data_rdtx, comorbidity, by= c("id"), all.x= TRUE)
data_rdtxc <- data_rdtxc %>% 
  mutate(transplantdate = as.Date(transplantdate, format="%d%b%Y"),
                 codate = as.Date(codate, format="%d%b%Y")) 

# Since there are 30 recordings for each patient - we take the recordings for when max(codate < tx_date) 
# i.e. the closest comorbidity recorded date (codate) to the transplant date
data_rdtxc <- data_rdtxc %>% 
  group_by(id, graftno, transplantdate) %>% 
  mutate(diff = transplantdate - codate) %>% 
  filter(diff>=0) %>%   # extract when transplant date after codate
  group_by(id, graftno) %>% 
  filter(diff == min(diff)) %>% # minimum difference between transplant and codate
  mutate(recip_waittime = as.numeric(as.character(recip_waittime)))

print(paste("data_rdtxc=",dim(data_rdtxc))) # 14973 x 101



```

# Adding the waitlist data

```{r}
# Extract only Active recipient dates
waitlist_data <- waitlist %>% dplyr::select(id, waitorgans, waitdate, waitstatus) %>% filter(waitstatus == "Active")

# Extracting the first "Active" date for the single transplant recipients 
waitdata <- waitlist_data[!duplicated(waitlist_data$id), ]

# Merge waitlist data with the dataset
data_rdtxcw <- merge(data_rdtxc,waitdata, by = "id" ) %>% mutate(waitdate = lubridate::dmy(waitdate))

print(paste("data_rdtxcw=",dim(data_rdtxcw))) # 7740 x 104



```



# Calculating epts and kdri

```{r}

index1 <- which(is.na(data_rdtxcw$ageattransplant) |
                             is.na(data_rdtxcw$diabetescode_text) | 
                             is.na(data_rdtxcw$recip_waittime))  

sub_data <- data_rdtxcw[-index1,]
length(index1)
nrow(sub_data)

data_rdtxcw[-index1,"recip_epts"]$recip_epts <-(0.047*as.numeric(sub_data$ageattransplant>25)*(sub_data$ageattransplant-25)
   -0.015*as.numeric(!(sub_data$diabetescode_text %in% c("No")))*as.numeric(sub_data$ageattransplant>25)*(sub_data$ageattransplant-25)
                  +0.398*(ifelse(sub_data$graftno == 1, 0, 1)) 
                  -0.237*as.numeric(!(sub_data$diabetescode_text %in% c("No")))*(ifelse(sub_data$graftno == 1, 0, 1)) 
                  +0.315*log(sub_data$recip_waittime/365.25+1) 
                  -0.099*as.numeric(!(sub_data$diabetescode_text %in% c("No")))*log(sub_data$recip_waittime/365.25+1)
                  +0.130*as.numeric(sub_data$recip_waittime<365.25)
                  -0.348*as.numeric(!(sub_data$diabetescode_text %in% c("No")))*as.numeric(sub_data$recip_waittime<365.25)
                  +1.262*as.numeric(!(sub_data$diabetescode_text %in% c("No"))))



index <- which(is.na(data_rdtxcw$donorage)|
                             is.na(data_rdtxcw$donor_height)|
                             is.na(data_rdtxcw$donor_weight)|
                             is.na(data_rdtxcw$donor_hypertensioncode) |
                             data_rdtxcw$donor_hypertensioncode == "Unknown"|
                             is.na(data_rdtxcw$donor_diabetescode) | 
                             is.na(data_rdtxcw$ddon_creatinineterminal) |
                             is.na(data_rdtxcw$ddon_deathcategorycode) | 
                            is.na(data_rdtxcw$ddon_heartbeatingcode))

sub_data <- data_rdtxcw[-index,]
length(index)
nrow(sub_data)


data_rdtxcw[-index,"donor_kdri"]$donor_kdri <- exp(0.0128*(sub_data$donorage-40)
                           -0.0194*as.numeric(sub_data$donorage<18)*(sub_data$donorage-18)
                           + 0.0107*as.numeric(sub_data$donorage>50)*(sub_data$donorage-50)
                           -0.0464*(sub_data$donor_height-170)/10
                           -0.0199*as.numeric(sub_data$donor_weight<80)*(sub_data$donor_weight-80)/5
                           +0.1260*as.numeric(sub_data$donor_hypertensioncode=="Yes")
                           +0.1300*as.numeric(!(sub_data$donor_diabetescode %in% c("No", "No Diabetes")))
                           +0.0881*as.numeric(sub_data$ddon_deathcategorycode %in% c("Cerebral Infarct", "Intracranial Haemorrhage"))
                           +0.2200*(sub_data$ddon_creatinineterminal/88.42-1)
      -0.2090*as.numeric(sub_data$ddon_creatinineterminal/88.42>1.5)*(sub_data$ddon_creatinineterminal/88.42-1.5)
      +0.1330*as.numeric(sub_data$ddon_heartbeatingcode == "N"))



  
summary(data_rdtxcw$donor_kdri)
summary(data_rdtxcw$recip_epts)


```



# Add Variable names 

```{r}
# Adding the variable names
data_name_dictionary <- read_excel("./42510_paeds_allocation_project/data_name_dictionary.xlsx")
index <- which(is.na(data_name_dictionary$new_name))

names(data_rdtxcw)[names(data_rdtxcw) %in% c(data_name_dictionary$raw_data_name[-index])] <- c(data_name_dictionary$new_name[-index])

```

# Fixing the recip_liststartdate for multiple deceased donor transplants

```{r}
# Extracting the second/third "Active" date for multiple transplant 
multi<-names(which(table(data_rdtxcw$recip_id) >= 2))
length(multi) #498 multiple transplants


# Extracting the recipients with multiple transplants from the main dataset and waitlist dataset
multi_transplant <-data_rdtxcw[data_rdtxcw$recip_id %in% multi,] %>% 
  dplyr::select(recip_id,tx_date, recip_graftno, recip_liststartdate) %>% 
  group_by(recip_id) %>% 
  arrange(recip_graftno, .by_group = TRUE)
waitdata <- waitlist_data[waitlist_data$id %in% multi,] %>% 
  mutate(waitdate = lubridate::dmy(waitdate))

# Merging datasets and extracting the first minimum eligible list date before the 2nd transplant but after the 1st transplant
waitdata <- merge(multi_transplant, waitdata, by.x = "recip_id", by.y = "id") %>% 
  mutate(diff = as.numeric(waitdate - tx_date)) %>% 
  group_by(recip_id) %>% 
  mutate(chosen = ifelse(diff >0, "eligible", "."))%>% 
  filter(chosen =="eligible") %>% 
  group_by(recip_id) %>% 
  mutate(newdate = min(waitdate)) # minimum time between the list date and the 1st transplant, but after the 1st transplant and before the 2nd. 

# Fixing the list start date for the multiple transplants
for(i in multi){
  data_rdtxcw[data_rdtxcw$recip_id == i & data_rdtxcw$recip_graftno ==2, "recip_liststartdate"] <- waitdata[waitdata$recip_id ==i, ]$newdate[1]
}

dim(data_rdtxcw) #7740 x 107

saveRDS(data_rdtxcw, file = "data_2017_full.rds")


```


# Extracting Deceased Donors, (keeping multiple transplant) and transplant and RRT year 2008/2006+

```{r }
## Extract the subset desired: deceased donors, kidney only, within certain time frame 

data_rdtxcw_final <- data_rdtxcw  %>%
  filter(donor_source == "Deceased", 
          tx_date > "2008-09-18", 
         is.na(tx_organ),
         recip_liststartdate > "2006-06-28", 
         recip_rrtstartdate > "2006-06-28")%>% 
  mutate(recip_listtime = as.numeric(tx_date - recip_liststartdate))%>% # calculate the recipient list time 
  as.data.frame()

data_rdtxcw_final <- data_rdtxcw_final %>% mutate_if(is.factor, as.character) %>% mutate(recip_id = as.character(recip_id)) 
data_rdtxcw_final$tx_organ <- "kidneys"

print(paste("data_rdtxcw_final=",dim(data_rdtxcw_final))) # 3981 x 107 
saveRDS(data_rdtxcw_final, "data_2017.rds") 


```


# Waitlist only 

```{r}
waitlist_active <- waitlist[waitlist$waitstatus == "Active" & is.na(waitlist$waitorgans),]
waitlist_active <- waitlist_active[!(waitlist_active$id %in% data_rdtxcw$recip_id),]
waitlist_active<- waitlist_active[!duplicated(waitlist_active$id),]

active_patients <- merge(waitlist_active, patients, by.x="id", by.y = "id")

active_patients$rrtstartdate<- lubridate::dmy(active_patients$rrtstartdate)
active_patients$waitdate<- lubridate::dmy(active_patients$waitdate)
max(data_rdtxcw$tx_date)


active_patients <- active_patients %>% filter(rrtstartdate > "2006-06-28", 
                                              waitdate > "2006-06-28" ) %>% 
  mutate(recip_waittime = as.numeric(as.Date("2017-12-31") - rrtstartdate)/365,
         recip_listtime = as.numeric(as.Date("2017-12-31") - waitdate)/365)


saveRDS(active_patients, "active_patients.rds") 




```